Refining from the BON High-level Specification Language to the VHDL-AMS Hardware Description Language

### 2009-08-27 papers/BON_VHDL/paper.tex
more ...

%======================================================================
\section{Introduction}
\label{sec:introduction}
The \ST sensor platform, represents heterogeneous collection of digital, analog 
and mixed signal hardware circuits. Embedded system complexity 
requires the need to strive for formal descriptions each component 
separatelly in order to produce specification of the \STS. 
Our aim is to be able to precisely describe the structures of circuits,
the behaviour of circuits and the formal link between structures and behaviours.
In other words, the behavioural specification of the \STS may be 
specified as a \inv on the signals at the circuit's ports and may be obtained 
by taking the conjunction of the \emph{invariants} that describe 
behaviours of the individual components separately. 
For this purpose we need  sophisticated tool and methodology 
in order to be able to achieve this goal.  
The authors will illustrate in this article how BON can represent a high 
level analogue mapping and will further illustrate a procedure for 
transformation from VHDL-AMS to BON.\\ 
This article provides two main contributions.
\begin{enumerate}
\item 
 It provides a solution how to formally emulate
 piece of hardware using BON-JML approach; up to date BON and JML
 has been used mostly in the pure object oriented software construction.    
\item
It relates hardware description code and formal calculus. 
This means that if there  is a code that is "correct by construction", 
then it will be derivable from BON (informal/formal) specification, 
via the refinement calculus to JML implementation.     
\end{enumerate}
The rest of this article is organized as follows.
In \autoref{sec:vhdlams} we have  presented an overview of VHDL-AMS language.
\autoref{sec:choiceobon} defines  BON principles and explains the 
reasons why we have decided to use BON method.
The central part of the our article is \autoref{sec:mappingVB}, which 
shows mapping between VHDL-AMS and BON. 
In \autoref{sec:refVB} we will give practical aspect of refinement, 
applied on real capacitor component, in order to produce formal specification  
and software simulation.\\  
A comprehensive list of the mapping between BON and VHDL-AMS can be 
found in the appendix.
%=====================================================================

@@

\subsection{Seamlessness and Reversibility}
\label{subsec:seamrever}
\emph{Seamlessness} is the use of a continuous process 
throughout the software lifecycle.
\emph{Reversibility} is the support for both forward and 
backward development process: from analysis to design 
and implementation, and back.
In VHDL-AMS, modelling represents an continious top-down-bottom-up
process. Top-down approach serves for design while bottom-up
serves for verification. The possibility of seamlessly translating 
changes made during a certain development phase back into earlier phases,
facilitates a stable and predictable environment and ensures the various 
models consistent with the code. Keeping in sync various models, documetation
and specification from forward progress to backward and redesign, preclude 
that design and implementation diverge to an inconsistent state. 
\emph{Seamlessness} also plays an important role in relating behaviours at
different levels of abstraction. For example, one may wish to analyze
a "real capacitor" at the level of individual components (a resistor and ideal capacitor)
and their behaviours, or at the more abstract level through general behaviour (\inv) and
using existential quantification (\texttt{exists}) to encapsulate all internal signals. 
Therefore capturing similarities between the differing levels of abstractions 
far outweigh their inevitable differences, making it possible to obtain a 
continuous process that ensures a direct mapping between a formal specification 
and its software solution.

\subsection{Contracting}
\label{subsec:contracting}
\emph{Contracting} is the definition, for each software element, 
of the associated semantic properties. BON use a full assertion mechanism, 
allowing analysts to specify not only the structure of a system but also its
semantics ( constraints and invariants of the expected results ). 
In VHDL-AMS \emph{simultaneous statement} represents analogue constraint of 
continuous-time waveform. In other words, analog kernel calculate quantities such 
that the constraints specified by the simultaneous statements are satisfied. 
We will agree that the constraint of waveform is nothing more than \inv.
Another example of \inv is conservation of energy law in which 
the summation of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) 
around any loop equals zero and the summation of through 
quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node 
(\texttt{terminal}) equals zero. Moreover, any predefined (implicite) attribute in VHDL-AMS  
has defined pre (post) condition trought input (output) 
parameters and theirs constraints. Furthermore any subtype (subnature) 
represents restricted set of values from a base type (nature). 
The condition that determines which values are in the subtype (subnature)  
may be contracted by specifying characteristic predicate which represents \inv.\\ 
Another advantage of BON is an easily-readable, English-based textual notation. 
The current IEEE standard VHDL-AMS language reference manual tries to define 
VHDL-AMS as well as possible in a descriptive way, explaining the 
semantics in English.\\
Therefore, on the basis of the above, BON is the obvious choice 
for the \STS.
Throughput this article we will make this more clearly.  
\section{Mapping between VHDL-AMS and BON}
\label{sec:mappingVB}
The main theme of this article is the mapping between VHDL-AMS and BON.
\section{Refinement from VHDL-AMS to BON }
\label{sec:refVB} 
An VHDL-AMS design units are described using an \ent-\arch structure. 
The design unit contains an \ent and one or more \archs. 
The \ent specifies the interface that an entity provides of itself 
to the outside. In other words, \ent introduces a name, list of 
input/output ports and the definition of the generic parameters. 
The textual presentation of \ent is shown in Figure.1. 
\input{adc-entity}

>>>>

%======================================================================
\section{Introduction}
\label{sec:introduction}
The \ST sensor platform, represents heterogeneous collection of digital, analog
and mixed signal hardware circuits. Embedded system complexity
requires the need to strive for formal descriptions each component
separately in order to produce specification of the \STS.
Our aim is to be able to precisely describe the structures of circuits,
the behavior of circuits and the formal link between structures and behaviors.
In other words, the behavioral specification of the \STS may be
represented as a \inv on the signals at the circuit's ports and may be obtained
by taking the conjunction of the \emph{invariants} that describe
behaviors of the individual components separately.
For this purpose we need  sophisticated tool and methodology
in order to be able to achieve this goal. 
The authors will illustrate in this article how BON can represent a high
level analogue mapping and will further illustrate a procedure for
transformation from VHDL-AMS to BON.\\
This article provides two main contributions.
\begin{enumerate}
\item
 It provides a solution how to formally emulate
 piece of hardware using BON-JML approach; up to date BON and JML
 have been used mostly in the pure object oriented software construction.   
\item
It relates hardware description code and formal calculus.
This means that if there  is a code that is "correct by construction",
then it will be derivable from BON (informal/formal) specification,
via the refinement calculus to JML implementation.    
\end{enumerate}
The rest of this article is organized as follows.
In \autoref{sec:vhdlams} we have  presented an overview of VHDL-AMS language.
\autoref{sec:choiceobon} defines  BON principles and explains the
reasons why we have decided to use BON method.
The central part of the our article is \autoref{sec:mappingVB}, which
shows mapping between VHDL-AMS and BON.
In \autoref{sec:refVB} we will give practical aspect of refinement,
applied on real capacitor component, in order to produce formal specification 
and software simulation.\\ 
A comprehensive list of the mapping between BON and VHDL-AMS can be
found in the appendix.
%=====================================================================

@@

\textbf{Seamlessness and  Reversibility} \xspace \xspace
Seamlessness is the use of a continuous process
throughout the software lifecycle. Reversibility is the support for
both forward and backward development process: from analysis to design
and implementation, and back.
In VHDL-AMS, modelling represents an continious top-down-bottom-up
process. Top-down approach serves for design while bottom-up
serves for verification. The possibility of seamlessly translating
changes made during a certain development phase back into earlier phases,
facilitates a stable and predictable environment and ensures the various
models consistent with the code. Keeping in sync various models, documetation
and specification from forward progress to backward and redesign, preclude
that design and implementation diverge to an inconsistent state.\\
\emph{Seamlessness} also plays an important role in relating behaviours at
different levels of abstraction. For example, one may wish to analyze
a "real capacitor" at the level of individual components (a resistor and
ideal capacitor) and their behaviours, or at the more abstract level through
general behaviour (\inv) and using existential quantification (\texttt{exists})
to encapsulate all internal structure. Therefore capturing similarities between
the differing levels of abstractions far outweigh their inevitable differences,
making it possible to obtain a continuous process that ensures a direct mapping
between a formal specification and its software solution.\\
\textbf{Contracting} \xspace \xspace Contracting is the definition, for each
software element, of the associated semantic properties. BON use a full
assertion mechanism, allowing analysts to specify not only the structure of
a system but also its semantics ( constraints and invariants of the expected results ).
The common way to specify a program statement is to give a \emph{precondition} and
\emph{postcondition} on the state that the implementation must satisfy, assuming that
only certain program variables are changed.
In VHDL-AMS any predefined (implicite) attribute 
has defined \emph{pre} (\emph{post}) \emph{condition} trought input (output)
parameters and theirs constraints. Furthermore, any subtype (subnature)
represents restricted set of values from a base type (nature).
The condition that determines which values are in the subtype (subnature) 
may be contracted by specifying characteristic predicate which represents \inv.\\
Let us next look how \emph{simultaneous statements} are handled in contract design.
The \emph{simultaneous statement} represents analogue constraint of
continuous-time waveform. We will assume that the constraint of waveform
represents \inv. This premise deserve an informal explanation.\\
Briefly stated, analog kernel calculate quantities such
that the constraints specified by the simultaneous statements are satisfied.
More accurately, \emph{simultaneous statements} can be thought of as
algebraic equations. It means that the number of simultaneous statements
should be equal to the number of unknowns, i.e. \emph{free} and \emph{branch}
\emph{quantities}. When we reasoning about conservation of energy law, for example,
the \emph{simultaneous statements} constrain the values of \emph{branch quantities}
in a model.
In other words, the sum of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0})
around any closed path must be equal to zero and the sum of \through
quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node
(terminal) must be equal to zero. To be more precise, the results of summations 
above remains zero when a particular transformation is applied to quantities.
In general, they are \inv. \\
However if our implementation does not satisfy conservation of energy law, 
 then the invariant will be broken.
In practice, one can use \emph{invariant} against the already existing code in
order to check the correctness.
It is important because we cannot patch a fabricated chip once company started
massive production. \\
Another advantage of BON is an easily-readable, English-based textual notation.
The current IEEE standard VHDL-AMS language reference manual tries to define
VHDL-AMS as well as possible in a descriptive way, explaining the
semantics in English.\\
Therefore, on the basis of the above, BON is the obvious choice
for the \STS.
Throughput this article we will make this more clearly. 
\section{Mapping between VHDL-AMS and BON}
\label{sec:mappingVB}
The main theme of this article is the mapping between VHDL-AMS and BON.
\section{Refinement from VHDL-AMS to BON }
\label{sec:refVB}
An VHDL-AMS design units are described using an \ent-\arch structure.
The design unit contains an \ent and one or more \archs.
The \ent specifies the interface that an entity provides of itself
to the outside. In other words, \ent introduces a name, list of
input/output ports and the definition of the generic parameters.
The textual presentation of \ent is shown in Figure.1.
\input{adc-entity}
